---
title: "Regression and Other Stories: Earnings"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predict respondents' yearly earnings using survey data from
1990. See Chapters 6, 9 and 12 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Seed
SEED <- 7783
  # Earnings data
file_earnings <- here::here("Earnings/data/earnings.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# Chapter 6

## Data

```{r, message=FALSE}
earnings <- 
  file_earnings %>% 
  read_csv() %>% 
  mutate(
    sex = 
      case_when(
        male == 0 ~ "female",
        male == 1 ~ "male",
        TRUE ~ NA_character_
      )
  )

earnings %>% 
  select(height, sex, earn)
```

## Linear regression of earnings on height and sex with no interaction

```{r}
fit_2 <- 
  stan_glm(earn ~ height + sex, data = earnings, seed = SEED, refresh = 0)

print(fit_2)
```

```{r}
fit_params <- 
  tribble(
    ~sex, ~intercept, ~slope,
    "female", coef(fit_2)[["(Intercept)"]], coef(fit_2)[["height"]],
    "male", 
      coef(fit_2)[["(Intercept)"]] + coef(fit_2)[["sexmale"]],
      coef(fit_2)[["height"]]
  )

offset <- 0.2

earnings %>% 
  mutate(
    height =
      case_when(
        sex == "female" ~ height - offset,
        sex == "male" ~ height + offset,
        TRUE ~ NA_real_
      )
  ) %>% 
  ggplot(aes(height, earn, color = sex)) +
  geom_count() +
  geom_abline(
    aes(slope = slope, intercept = intercept, color = sex),
    data = fit_params
  ) +
  coord_cartesian(ylim = c(0, 1e5)) +
  scale_x_continuous(breaks = scales::breaks_width(1), minor_breaks = NULL) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_color_discrete(labels = c("Female", "Male")) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Linear regression of earnings on height and sex with no interaction",
    x = "Height",
    y = "Earnings",
    color = "Sex",
    size = "Count"
  )
```

The equations for the regression lines are:

    Men:   y = `r format(fit_params$intercept[fit_params$sex == "male"], digits = 0, nsmall = 0, scientific = FALSE)` + `r format(fit_params$slope[fit_params$sex == "male"], digits = 0, nsmall = 0, scientific = FALSE)` x
    Women: y = `r format(fit_params$intercept[fit_params$sex == "female"], digits = 0, nsmall = 0, scientific = FALSE)` + `r format(fit_params$slope[fit_params$sex == "female"], digits = 0, nsmall = 0, scientific = FALSE)` x

## Linear regression of earnings on height and sex with interaction

```{r}
fit_3 <- 
  stan_glm(
    earn ~ height + sex + height:sex,
    data = earnings,
    seed = SEED,
    refresh = 0
  )

print(fit_3)
```

```{r}
fit_params <- 
  tribble(
    ~sex, ~intercept, ~slope,
    "female", coef(fit_3)[["(Intercept)"]], coef(fit_3)[["height"]],
    "male", 
      coef(fit_3)[["(Intercept)"]] + coef(fit_3)[["sexmale"]],
      coef(fit_3)[["height"]] + coef(fit_3)[["height:sexmale"]]
  )

offset <- 0.2

earnings %>% 
  mutate(
    height =
      case_when(
        sex == "female" ~ height - offset,
        sex == "male" ~ height + offset,
        TRUE ~ NA_real_
      )
  ) %>% 
  ggplot(aes(height, earn, color = sex)) +
  geom_count() +
  geom_abline(aes(slope = slope, intercept = intercept, color = sex), data = fit_params) +
  coord_cartesian(ylim = c(0, 1e5)) +
  scale_x_continuous(breaks = scales::breaks_width(1), minor_breaks = NULL) +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_color_discrete(labels = c("Female", "Male")) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Linear regression of earnings on height and sex with interaction",
    x = "Height",
    y = "Earnings",
    color = "Sex",
    size = "Count"
  )
```

The equations for the regression lines are:

    Men:   y = `r format(fit_params$intercept[fit_params$sex == "male"], digits = 0, nsmall = 0, scientific = FALSE)` + `r format(fit_params$slope[fit_params$sex == "male"], digits = 0, nsmall = 0, scientific = FALSE)` x
    Women: y =  `r format(fit_params$intercept[fit_params$sex == "female"], digits = 0, nsmall = 0, scientific = FALSE)` + `r format(fit_params$slope[fit_params$sex == "female"], digits = 0, nsmall = 0, scientific = FALSE)` x

From the plots, we can see that many more women than men have no earnings.

```{r}
earnings %>% 
  count(earn, sex) %>% 
  group_by(sex) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup() %>% 
  filter(earn == 0)
```

15% of women have no earnings, whereas only 2% of men have no earnings.

