---
title: "Regression and Other Stories: Height and weight"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Predict weight from height. See Chapters 9 and 10 in Regression
and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Earnings data
file_earnings <- here::here("Earnings/data/earnings.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# Chapter 9

## Data

```{r, message=FALSE}
earnings <- 
  file_earnings %>% 
  read_csv() %>% 
  mutate(
    sex = 
      case_when(
        male == 0 ~ "Female",
        male == 1 ~ "Male",
        TRUE ~ NA_character_
      )
  )

earnings %>% 
  select(height, weight, sex, ethnicity)
```

```{r}
mean(earnings$height)
```

## Simulating uncertainty for the linear predictor and new observations

Fit linear regression of weight against height.

The option `refresh = 0` suppresses the default Stan sampling progress output. This is useful for small data with fast computation. For more complex models and bigger data, it can be useful to see the progress.

```{r}
set.seed(733)

fit_1 <- stan_glm(weight ~ height, data = earnings, refresh = 0)

fit_1
```

Predict the weight for someone 66 inches tall.

```{r}
set.seed(733)

pred_66 <- posterior_predict(fit_1, newdata = tibble(height = 66))
```

```{r}
cat(
  str_glue(
    "The predicted weight for someone 66 inches tall is ",
    "{format(mean(pred_66), digits = 1, nsmall = 1)} pounds with a sd of ",
    "{format(sd(pred_66), digits = 1, nsmall = 1)}."
  )
)
```

Center heights.

```{r}
earnings <- 
  earnings %>% 
  mutate(c_height = height - 66)
```

Fit using centered heights.

```{r}
set.seed(733)

fit_2 <- stan_glm(weight ~ c_height, data = earnings, refresh = 0)

fit_2
```

New data for someone 70 inches tall.

```{r}
new <- tibble(c_height = 4)
```

Point prediction.

```{r}
point_pred <- predict(fit_2, newdata = new)

as.double(point_pred)
```

Posterior simulations where variation is from posterior uncertainty in the coefficients.

```{r}
linpred <- posterior_linpred(fit_2, newdata = new)
```

Posterior simulations where variation is from posterior uncertainty in the coefficients and predictive uncertainty.

```{r}
set.seed(733)

pred <- posterior_predict(fit_2, newdata = new)
```

```{r}
preds <- tibble(linpred = as.double(linpred), pred = as.double(pred))

preds %>% 
  summarize(across(everything(), list(median = median, mad = mad)))
```

The medians of the two sets of posterior simulations are similar, but the median absolute deviations are very different.

### Plot

```{r}
preds %>% 
  ggplot(aes(y = stat(density))) +
  geom_freqpoly(aes(pred, color = "pred"), binwidth = 10) +
  geom_freqpoly(aes(linpred, color = "linpred"), binwidth = 0.2) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  scale_y_continuous(breaks = 0) +
  scale_color_discrete(
    breaks = c("linpred", "pred"),
    labels = c("Linear predictor", "Predicted value")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Distributions of posterior predictions",
    x = "Weight (pounds)",
    y = NULL,
    color = NULL
  )
```

