---
title: "Regression and Other Stories: Arsenic"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Building a logistic regression model: wells in Bangladesh. See
Chapters 13 and 14 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Data on arsenic in wells in Bangladesh
file_wells_all <- here::here("Arsenic/data/all.dta")
  # Data on arsenic in wells in Bangladesh that exceed the safety standard
file_wells <- here::here("Arsenic/data/wells.csv")
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

# 13 Logistic regression

## 13.7 Building a logistic regression model: wells in Bangladesh

### Background

#### Data on all wells

Data on arsenic level in all wells and their location.

```{r}
wells_all <- 
  file_wells_all %>% 
  haven::read_dta() %>% 
  transmute(arsenic = as / 100, x, y)

summary(wells_all)
```

The position data appear to have some problems.

```{r}
sort(wells_all$x) %>% 
  head(100)

sort(wells_all$y) %>% 
  head(100)
```

Well remove the problematic rows and center and scale the positions.

```{r}
wells_all <- 
  wells_all %>% 
  drop_na(arsenic) %>% 
  filter(x > 91, y > 24) %>% 
  mutate(
    x = (x - mean(x)) / 1000,
    y = (y - mean(y)) / 1000
  )
```

Distribution of arsenic level in all wells.

```{r}
wells_all %>% 
  ggplot(aes(arsenic)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  geom_vline(xintercept = 0.5, color = "red") +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Distribution of arsenic level in all wells",
    subtitle = "Levels lower than the vertical line are considered safe",
    x = "Arsenic level",
    y = "Count"
  )
```

```{r}
quantile(wells_all$arsenic, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1))
```

There are some outliers with levels up to 9.65, but the majority of the wells have levels under 3.

```{r}
v <- mean(wells_all$arsenic > 0.5)
v
```

About `r format(100*v, digits = 1, nsmall = 1)`% of wells are unsafe; that is, have arsenic levels over 0.5.

Map of wells in an area of Araihazar, Bangladesh.

```{r, fig.asp=1.25}
wells_all %>% 
  mutate(
    type = if_else(arsenic > 0.5, "Unsafe wells", "Safe wells"),
    arsenic = if_else(arsenic > 3, 3, arsenic)
  ) %>% 
  ggplot(aes(x, y, color = arsenic)) +
  geom_point(alpha = 0.75, size = 0.1) +
  coord_fixed() +
  facet_grid(rows = vars(type)) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  scale_y_continuous(breaks = scales::breaks_width(1)) +
  scale_color_viridis_c(breaks = c(0.5, 1:3), labels = c(c(0.5, 1:2), "3+")) +
  labs(
    title = "Map of wells in an area of Araihazar, Bangladesh",
    x = NULL,
    y = NULL,
    color = "Arsenic\nlevel"
  )
```

It looks as though the wells may be near dwellings along streets. In many areas unsafe wells are close to safe wells.

#### Data on wells that exceed the safety standard

Data.

```{r, message=FALSE}
wells <- read_csv(file_wells)

summary(wells)
```

The variables we will use here are:

* `switch`: Outcome variable:
  * 1 if household switched to a new well
  * 0 if household continued using its own well
* `arsenic`: Arsenic level of respondent's well
* `dist`: Distance (in meters) to the closest known safe well
* `dist100` = `dist / 100`

##### 1D EDA

Whether household switched.

```{r}
wells %>% 
  count(switch) %>% 
  mutate(prop = n / sum(n))
```

About 57.5% of households switched.

Distribution of arsenic level in unsafe wells.

```{r}
wells %>% 
  ggplot(aes(arsenic)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Distribution of arsenic level in unsafe wells",
    x = "Arsenic level",
    y = "Count"
  )
```

Distance to the closest known safe well.

```{r}
wells %>% 
  ggplot(aes(dist)) +
  geom_histogram(binwidth = 10, boundary = 0) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Distance to the closest known safe well",
    x = "Distance (meters)",
    y = "Count"
  )
```

```{r}
quantile(wells$dist, probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1))
```

The median distance to a safe well is about 37 meters. About 90% of the households have a safe well within 100 meters.

##### 2D EDA

Percentage of households who switched to new well by arsenic level.

```{r, message=FALSE}
wells %>% 
  ggplot(aes(arsenic, switch)) +
  stat_ydensity(
    aes(group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_smooth() +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = 
      "Percentage of households who switched to new well by arsenic level",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Arsenic level",
    y = "Percentage of households who switched"
  )
```

As expected, the percentage of households increases with the arsenic level in their well, from about 40% for wells that are just over the safety threshold to perhaps 80% for very high levels. The sparse data for high arsenic levels results in a large uncertainty.

Percentage of households who switched to new well by distance.

```{r, message=FALSE}
wells %>% 
  ggplot(aes(dist, switch)) +
  stat_ydensity(
    aes(group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_smooth() +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.1),
    minor_breaks = NULL,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Percentage of households who switched to new well by distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Percentage of households who switched"
  )
```

As expected, the percentage of households decreases with the distance to the closest known safe well, from about 60% when a safe well is very close to perhaps 20% when a safe well is far away. The sparse data for large distances results in a large uncertainty.

### Logistic regression with just one predictor

Fit a model using distance to the nearest safe well.

```{r}
set.seed(733)

fit_1 <- 
  stan_glm(
    switch ~ dist,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

print(fit_1, digits = 3)
```

LOO log score

```{r}
loo_1 <- loo(fit_1)

loo_1
```

Fit a model using scaled distance to the nearest safe well.

```{r}
set.seed(733)

fit_2 <- 
  stan_glm(
    switch ~ dist100,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

fit_2
```

LOO log score

```{r}
loo_2 <- loo(fit_2)

loo_2
```

### Graphing the fitted model

Probability of household switching to new well by distance.

```{r}
v <- 
  tibble(
    dist = seq_range(wells$dist),
    .pred = predict(fit_1, type = "response", newdata = tibble(dist))
  )

wells %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred), data = v) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Probability of household switching to new well by distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching"
  )
```

The empirical plot in the 2D EDA section showed that about 60% of households switched at the closest distance and about 20% switched at the farthest distance. The probabilities of the model at the extremes are similar. Between the distance of 0 - 50 meters, the empirical plot has a roughly constant percentage of households who switched, whereas the model has a linear decline.

### Interpreting the logistic regression coefficients

Proportion of households who switched.

```{r}
switch_prop <- mean(wells$switch)

switch_prop
```

Log score for model with constant prediction equal to the proportion of households who switched.

```{r}
log(switch_prop) * sum(wells$switch) + log(1 - switch_prop) * sum(!wells$switch)
```

Since the LOO log score for the model with distance is -2040, distance supplies some predictive information.

### Adding a second input variable

Fit a model using scaled distance and arsenic level.

```{r}
set.seed(733)

fit_3 <- 
  stan_glm(
    switch ~ dist100 + arsenic,
    family = binomial(link = "logit"),
    data = wells,
    refresh = 0
  )

fit_3
```

LOO log score

```{r}
loo_3 <- loo(fit_3)

loo_3
```

Compare log scores.

```{r}
loo_compare(loo_2, loo_3)
```

Model 3 has the better log score, indicating that including arsenic level in the model clearly improves the predictive accuracy.

### Graphing the fitted model with two predictors

Probability of household switching to new well by distance and arsenic level.

```{r, fig.asp=0.75}
v <- 
  tibble(
    arsenic = c(0.5, quantile(wells$arsenic, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(arsenic) == "" ~ as.character(arsenic),
        TRUE ~ 
          str_glue(
            "{format(arsenic, digits = 1, nsmall = 1)} ({names(arsenic)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    dist = list(seq_range(wells$dist))
  ) %>% 
  unnest(dist) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_3, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by distance and arsenic level",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching",
    color = "Arsenic level (Quantile)"
  )
```

The probability decreases with distance and increases with arsenic level.

Probability of household switching to new well by arsenic level and distance.

```{r, fig.asp=0.75}
v <- 
  tibble(
    dist = c(0, quantile(wells$dist, probs = c(0.25, 0.5, 0.75))),
    label = 
      case_when(
        names(dist) == "" ~ as.character(dist),
        TRUE ~ 
          str_glue(
            "{format(dist, digits = 0, nsmall = 0)} ({names(dist)})"
          ) %>% 
          as.character()
      ) %>% 
      fct_inorder(),
    arsenic = list(seq_range(wells$arsenic))
  ) %>% 
  unnest(arsenic) %>% 
  mutate(
    dist100 = dist / 100,
    .pred = 
      predict(fit_3, type = "response", newdata = tibble(arsenic, dist100))
  )

v %>% 
  ggplot(aes(arsenic)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred, color = label)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  theme(legend.position = "bottom") +
  labs(
    title = 
      "Probability of household switching to new well by arsenic level and distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Arsenic level",
    y = "Probability of household switching",
    color = "Distance in meters (Quantile)"
  )
```

The probability increases with arsenic level and decreases with distance.

# 14 Working with logistic regression

## 14.3 Predictive simulation

Probability of household switching to new well by distance with uncertainty.

```{r}
new <- tibble(dist = seq_range(wells$dist))
linpred <- posterior_linpred(fit_1, newdata = new)
v <- 
  new %>% 
  mutate(
    .pred = predict(fit_1, type = "response", newdata = new),
    `5%`  = apply(linpred, 2, quantile, probs = 0.05) %>% plogis(),
    `25%` = apply(linpred, 2, quantile, probs = 0.25) %>% plogis(),
    `75%` = apply(linpred, 2, quantile, probs = 0.75) %>% plogis(),
    `95%` = apply(linpred, 2, quantile, probs = 0.95) %>% plogis()
  )

v %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    data = wells,
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), alpha = 0.25) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), alpha = 0.5) +
  geom_line(aes(y = .pred)) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Probability of household switching to new well by distance",
    subtitle =
      "With 50% and 90% predictive intervals\nVoilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching"
  )
```

In the region of sparse data for large distances, the uncertainty from the posterior distribution of the model is much less than the uncertainty seen in the 2D EDA section using LOESS.

