---
title: "Regression and Other Stories: SimpleCausal"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: readable
    toc: true
    toc_depth: 2
    toc_float: true
    code_download: true
---
Tidyverse version by Bill Behrman.

Simple graphs illustrating regression for causal inference. See
Chapter 1 in Regression and Other Stories.

The simulated data depends on the random seed, and thus the plots
and numbers here and in the book may differ. You can experiment
with the simulation variation by changing the seed.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(arm)

# Parameters
  # Seed for random numbers
SEED <- 1151
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

## Simulated data from linear model

```{r}
set.seed(SEED)

n <- 50
df_1 <- 
  tibble(
    x = runif(n, min = 1, max = 5),
    x_binary = if_else(x < 3, 0, 1),
    y = rnorm(n, mean = 10 + 3 * x, sd = 3)
  )
```

### Regression with binary predictor

```{r}
lm_1a <- lm(y ~ x_binary, data = df_1)

display(lm_1a)
```

```{r}
df_1 %>% 
  ggplot(aes(x_binary, y)) +
  geom_point() +
  geom_abline(
    slope = coef(lm_1a)[["x_binary"]],
    intercept = coef(lm_1a)[["(Intercept)"]]
  ) +
  scale_x_continuous(
    breaks = 0:1,
    minor_breaks = NULL,
    labels = c("Control", "Treatment")
  ) +
  labs(
    title = "Regression with binary treatment",
    x = NULL,
    y = "Outcome measurement"
  )
```

### Regression with continuous predictor

```{r}
lm_1b <- lm(y ~ x, data = df_1)

display(lm_1b)
```

```{r}
label = 
  str_glue(
    "Estimated treatment\n",
    "effect per unit of x is\n",
    "slope of fitted line: {format(coef(lm_1b)[['x']], digits = 1, nsmall = 1)}"
  )

df_1 %>% 
  ggplot(aes(x, y)) +
  geom_point() +
  geom_abline(
    slope = coef(lm_1b)[["x"]],
    intercept = coef(lm_1b)[["(Intercept)"]]
  ) +
  annotate(
    "text",
    x = 1.5,
    y = 24,
    label = label,
    hjust = 0
  ) +
  labs(
    title = "Regression with continuous treatment",
    x = "Treatment level",
    y = "Outcome measurement"
  )
```

## Simulated data from nonlinear model

```{r}
set.seed(SEED)

n <- 50
df_2 <- 
  tibble(
    x = df_1$x,
    y_mean = 5 + 30 * exp(-x),
    y = rnorm(n, mean = y_mean, sd = 2)
  )
```


### Regression with continuous predictor

```{r}
lm_2 <- lm(y ~ x, data = df_2)

display(lm_2)
```

```{r}
df_2 %>% 
  ggplot(aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = y_mean)) +
  scale_y_continuous(breaks = scales::breaks_width(2)) +
  labs(
    title = "Nonlinear treatment effect",
    x = "Treatment level",
    y = "Outcome measurement"
  )
```

```{r}
df_2 %>% 
  ggplot(aes(x)) +
  geom_point(aes(y = y)) +
  geom_abline(
    slope = coef(lm_2)[["x"]],
    intercept = coef(lm_2)[["(Intercept)"]]
  ) +
  scale_y_continuous(breaks = scales::breaks_width(2)) +
  labs(
    title = "Nonlinear treatment effect",
    x = "Treatment level",
    y = "Outcome measurement"
  )
```

## Simulated data from two groups

```{r}
set.seed(SEED)

n <- 100
not_used <- rnorm(n)  # So df_3 will match example
df_3 <- 
  tibble(
    x_2 = rep(0:1, n / 2),
    x_1 = 
      if_else(
        x_2 == 0,
        rnorm(n, mean = 0, sd = 1.2)^2,
        rnorm(n, mean = 0, sd = 0.8)^2
      ),
    y = rnorm(n, mean = 20 + 5 * x_1 + 10 * x_2, sd = 3)
  )
```

```{r}
lm_3 <- lm(y ~ x_1 + x_2, data = df_3)

display(lm_3)
```

```{r}
slope <- coef(lm_3)[["x_1"]]
intercept_0 <- coef(lm_3)[["(Intercept)"]]
intercept_1 <- coef(lm_3)[["(Intercept)"]] + coef(lm_3)[["x_2"]]

lines <- 
  tribble(
    ~slope,  ~intercept,
     slope, intercept_0,
     slope, intercept_1
  )

label <- 
  str_glue(
    "Estimated treatment\n",
    "effect is {format(coef(lm_3)[['x_2']], digits = 1, nsmall = 1)}"
  )

ggplot() +
  geom_point(aes(x_1, y, color = as.factor(x_2)), data = df_3) +
  geom_abline(aes(slope = slope, intercept = intercept), data = lines) +
  annotate(
    "segment",
    x = 4.2,
    xend = 4.2,
    y = intercept_0 + slope * 4.2,
    yend = intercept_1 + slope * 4.2,
    arrow = arrow(length = unit(0.04, units = "npc"), ends = "both")
  ) +
  annotate(
    "text",
    x = 4.2,
    y = 37,
    label = label,
    hjust = 0
  ) +
  scale_color_discrete(
    breaks = 0:1,
    labels = c("Controls", "Treated")
  ) +
  theme(legend.position = "bottom") +
  labs(
    title = "Continuous pre-treatment predictor and binary treatment",
    x = "Pre-treatment predictor",
    y = "Outcome measurement",
    color = NULL
  )
```
