---
title: "Regression and Other Stories: Fake midterm and final exam"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Fake dataset of 1000 students' scores on a midterm and final
exam. See Chapter 6 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

## Data

```{r}
set.seed(2243)

n_sims <- 1000

exams <- 
  tibble(
    true_ability = rnorm(n_sims, mean = 50, sd = 10),
    noise_1 = rnorm(n_sims, mean = 0, sd = 10),
    noise_2 = rnorm(n_sims, mean = 0, sd = 10),
    midterm = true_ability + noise_1,
    final = true_ability + noise_2
  )
```

## Linear regression

The option `refresh = 0` suppresses the default Stan sampling progress output. This is useful for small data with fast computation. For more complex models and bigger data, it can be useful to see the progress.

```{r}
fit <- stan_glm(final ~ midterm, data = exams, refresh = 0)

print(fit, digits = 2)
```

## Plot

```{r, fig.asp=1}
intercept <- coef(fit)[["(Intercept)"]]
slope <- coef(fit)[["midterm"]]

exams %>% 
  ggplot(aes(midterm, final)) +
  geom_point(size = 0.5) +
  geom_abline(slope = slope, intercept = intercept) +
  coord_fixed(xlim = c(0, 100), ylim = c(0, 100)) +
  labs(
    title = "Simulated midterm and final exam scores",
    x = "Midterm exam score",
    y = "Final exam score"
  )
```

