---
title: "Regression and Other Stories: Elections Economy"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    md_extension: +gfm_auto_identifiers
    toc: true
---
Tidyverse version by Bill Behrman.

Predicting presidential vote share from the economy. See Chapters
1, 7, 8, 9, and 22 in Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)
library(rstanarm)

# Parameters
  # U.S. Presidential election results and GDP growth
file_hibbs <- here::here("ElectionsEconomy/data/hibbs.dat")
  # Common code
file_common <- here::here("_common.R")
  
#===============================================================================

# Run common code
source(file_common)
```

# 1 Overview

## 1.2 Why learn regression?

Data.

```{r}
hibbs <- 
  file_hibbs %>% 
  read.table(header = TRUE) %>% 
  as_tibble()

hibbs
```

Plot.

```{r}
hibbs %>% 
  ggplot(aes(growth, vote)) +
  geom_hline(yintercept = 50, color = "grey60") +
  geom_text(aes(label = year)) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Forecasting the election from the economy",
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

Linear regression.

The option `refresh = 0` suppresses the default Stan sampling
progress output. This is useful for small data with fast
computation. For more complex models and bigger data, it can be
useful to see the progress.

```{r}
set.seed(264)

fit <- stan_glm(vote ~ growth, data = hibbs, refresh = 0)
```

Print default summary of the fitted model.

```{r}
fit
```

Print summary of the priors used.

```{r}
prior_summary(fit)
```

Almost all models in Regression and Other Stories have very good
sampling behavior. `summary()` function can be used to obtain the
summary of the convergence diagnostics for MCMC sampling.

```{r}
summary(fit)
```

Posterior interval.

```{r}
posterior_interval(fit) %>% 
  round(digits = 1)
```

Plot regression line.

```{r}
intercept <- coef(fit)[["(Intercept)"]]
slope <- coef(fit)[["growth"]]
eqn <- 
  str_glue(
    "y = {format(intercept, digits = 1, nsmall = 1)} + ",
    "{format(slope, digits = 1, nsmall = 1)} x"
  )

hibbs %>% 
  ggplot(aes(growth, vote)) +
  geom_hline(yintercept = 50, color = "grey60") +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point() +
  annotate("text", x = 3.05, y = 53.75, label = eqn, hjust = 0) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Data and linear fit",
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

# Chapter 7

## Plot of economy and election results

```{r, fig.asp=1}
v <- 
  hibbs %>% 
  mutate(
    label = str_glue("  {inc_party_candidate} vs. {other_candidate} ({year})  "),
    hjust = 
      if_else(
        year %in% c(1960, 1964, 1972, 1976, 1984, 1996, 2012),
        "right",
        "left"
      ),
    vjust =
      case_when(
        year == 1992 ~ 0,
        year == 2000 ~ 0.2,
        year %in% c(1960, 2008) ~ 1,
        TRUE ~ 0.5
      )
  )

v %>% 
  ggplot(aes(growth, vote)) +
  geom_hline(yintercept = 50, color = "grey60") +
  geom_point() +
  geom_text(aes(label = label, hjust = hjust, vjust = vjust)) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1, scale = 1),
    expand = expansion(add = c(0.8, 1.2))
  ) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Forecasting elections from the economy",
    x = "Income growth",
    y = "Incumbent party's share of the popular vote"
  )
```

```{r}
growth <- 2
```

## Plot of prediction given `r growth`% growth

```{r}
vote_pred <- predict(fit, newdata = tibble(growth = growth))
sd <- sigma(fit)

v <- 
  tibble(
    vote = c(50, seq(vote_pred - 4 * sd, vote_pred + 4 * sd, length.out = 201)),
    y = dnorm(vote, mean = vote_pred, sd = sd)
  )

label <-
  str_glue(
    "Predicted\n",
    "{format(100 * (1 - pnorm(50, mean = vote_pred, sd = sd)), digits = 0)}% ",
    "chance\n",
    "of Clinton victory"
  )

v %>% 
  ggplot(aes(vote, y)) +
  geom_area(data = . %>% filter(vote >= 50), fill = "grey70") +
  geom_line() +
  geom_segment(
    aes(x = 50, xend = 50, y = 0, yend = dnorm(50, mean = vote_pred, sd = sd))
  ) +
  annotate("text", x = 51.5, y = 0.025, label = label, hjust = 0) +
  scale_x_continuous(
    breaks = scales::breaks_width(5),
    labels = scales::label_percent(accuracy = 1, scale = 1)
  ) +
  scale_y_continuous(breaks = 0) +
  labs(
    title = "Probability forecast of Hillary Clinton vote share in 2016",
    subtitle = str_glue("Based upon {growth}% rate of economic growth"),
    x = "Clinton share of the two-party vote",
    y = NULL
  )
```

# Chapter 8

## Data and linear fit

```{r}
hibbs %>% 
  ggplot(aes(growth, vote)) +
  geom_abline(slope = slope, intercept = intercept) +
  geom_point() +
  annotate("text", x = 3.05, y = 53.75, label = eqn, hjust = 0) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Data and linear fit",
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

## Data and range of possible linear fits

```{r}
set.seed(189)

sims <- as_tibble(fit)
```

```{r}
set.seed(189)

n_lines <- 50

hibbs %>% 
  ggplot(aes(growth, vote)) +
  geom_abline(
    aes(slope = growth, intercept = `(Intercept)`),
    data = sims %>% slice_sample(n = n_lines),
    alpha = 0.25
  ) +
  geom_abline(slope = slope, intercept = intercept, color = "red") +
  geom_point(color = "white", size = 2) +
  geom_point() +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Data and range of possible linear fits",
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

# Chapter 9

## Propogating uncertainty in inference using posterior simulations

Fit parameter statistics.

```{r}
print(fit, digits = 2)
```

Extract the simulations.

```{r}
set.seed(189)

sims <- as_tibble(fit)
```

Median and mean absolute deviation of parameters.

```{r}
sims %>% 
  pivot_longer(
    cols = everything(),
    names_to = "param"
  ) %>% 
  group_by(param) %>% 
  summarize(
    median = median(value),
    mad = mad(value)
  )
```

Rename intercept and slope.

```{r}
sims <- 
  sims %>% 
  rename(a = "(Intercept)", b = "growth")
```

Median and mean absolute deviation for a derived quantity a / b.

```{r}
sims %>% 
  mutate(z = a / b) %>% 
  summarize(across(z, list(median = median, mad = mad)))
```

### Posterior simulations of the intercept

```{r}
a_median <- median(sims$a)
a_mad <- mad(sims$a)
lines_mad <- a_median + setdiff(-2:2, 0) * a_mad

sims %>% 
  ggplot(aes(a)) +
  geom_histogram(binwidth = 0.5, boundary = 0) +
  geom_vline(xintercept = a_median, color = "red") +
  geom_vline(xintercept = lines_mad, color = "grey60") +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  labs(
    title = "Posterior simulations of the intercept",
    subtitle = 
      "With posterior median in red and ± 1 and 2 median absolute deviations in grey"
  )
```

### Posterior simulations of the slope

```{r}
b_median <- median(sims$b)
b_mad <- mad(sims$b)
lines_mad <- b_median + setdiff(-2:2, 0) * b_mad

sims %>% 
  ggplot(aes(b)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  geom_vline(xintercept = b_median, color = "red") +
  geom_vline(xintercept = lines_mad, color = "grey60") +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Posterior simulations of the slope",
    subtitle = 
      "With posterior median in red and ± 1 and 2 median absolute deviations in grey"
  )
```

### Posterior draws of the regression coefficients a, b

```{r}
a <- coef(fit)[["(Intercept)"]]
b <- coef(fit)[["growth"]]

sims %>% 
  ggplot(aes(a, b)) +
  geom_point(size = 0.1) +
  geom_point(data = tibble(a = a, b = b), color = "red", size = 1.5) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  labs(title = "Posterior draws of the regression coefficients a, b")
```

```{r}
n_lines <- 100
```

### Data and `r n_lines` posterior draws of the line, y = a + b x

```{r}
set.seed(189)

hibbs %>% 
  ggplot(aes(growth, vote)) +
  geom_abline(
    aes(slope = b, intercept = a),
    data = sims %>% slice_sample(n = n_lines),
    alpha = 0.25
  ) +
  geom_abline(slope = b, intercept = a, color = "red") +
  geom_point(color = "white", size = 2) +
  geom_point() +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = 
      str_glue("Data and {n_lines} posterior draws of the line, y = a + b x"),
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

### Data and linear fit with 50% and 90% predictive intervals

```{r}
v <- 
  tibble(growth = seq_range(hibbs$growth)) %>% 
  predictive_intervals(fit = fit)

v %>% 
  ggplot(aes(growth)) +
  geom_ribbon(aes(ymin = `5%`, ymax = `95%`), alpha = 0.25) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), alpha = 0.5) +
  geom_line(aes(y = .pred)) +
  geom_point(aes(y = vote), data = hibbs, color = "white", size = 2) +
  geom_point(aes(y = vote), data = hibbs) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1, scale = 1)) +
  labs(
    title = "Data and linear fit with 50% and 90% predictive intervals",
    x = "Average recent growth in personal income",
    y = "Incumbent party's vote share"
  )
```

## Prediction and uncertainty

### Point prediction using `predict()`

Point prediction given 2% growth.

```{r}
new <- tibble(growth = 2)

y_point_pred <- predict(fit, newdata = new)
```

Manual calculation for point prediction.

```{r}
y_point_pred_manual <- mean(sims$a + sims$b * new$growth)

y_point_pred_manual - as.double(y_point_pred)
```

### Linear predictor with uncertainty using `posterior_linpred()`

Create a linear prediction for each of the `r nrow(sims)` posterior simulations of the model parameters.

```{r}
y_linpred <- posterior_linpred(fit, newdata = new)
```

Manual calculation for linear predictions.

```{r}
y_linpred_manual <- sims$a + sims$b * new$growth

max(abs(y_linpred_manual - y_linpred))
```

### Predictive distribution for a new observation using `posterior_predict()`

Predictive uncertainty.

```{r}
set.seed(835)

y_pred <- posterior_predict(fit, newdata = new)
```

Manual calculation for predictive uncertainty.

```{r}
set.seed(835)

y_pred_manual <- 
  sims$a + sims$b * new$growth + rnorm(nrow(sims), mean = 0, sd = sims$sigma)

max(abs(y_pred_manual - y_pred))
```

Plot of predictions.

```{r}
tibble(y_pred = as.numeric(y_pred)) %>% 
  ggplot(aes(y_pred)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = median(y_pred), color = "red") +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  labs(
    title = 
      str_glue(
        "Predicted percentage of two-party vote with {new$growth}% growth"
      ),
    subtitle = 
      str_glue(
        "Median prediction is {format(median(y_pred), digits = 1, nsmall = 1)}%"
      ),
    x = "Percentage of two-party vote",
    y = "Count"
  )
```

Summary of predictions.

```{r}
cat(
  str_glue(
    "Predicted Clinton percentage of two-party vote: ",
    "{format(median(y_pred), digits = 1, nsmall = 1)} with s.e. ",
    "{format(mad(y_pred), digits = 1, nsmall = 1)}\n",
    "Pr(Clinton win) = {format(mean(y_pred > 50), digits = 2, nsmall = 2)}"
  )
)
```

### Prediction given a range of input values

```{r}
new_grid <- tibble(growth = seq(-2, 4, 0.5))
y_point_pred_grid <- predict(fit, newdata = new_grid)
y_linpred_grid <- posterior_linpred(fit, newdata = new_grid)
y_pred_grid <- posterior_predict(fit, newdata = new_grid)
```

### Propogating uncertainty

```{r}
set.seed(672)

v <- 
  tibble(
    x_new = rnorm(nrow(sims), mean = 2, sd = 0.3),
    y_pred = rnorm(nrow(sims), mean = sims$a + sims$b * x_new, sd = sims$sigma)
  )

cat(
  str_glue(
    "Predicted Clinton percentage of two-party vote: ",
    "{format(median(v$y_pred), digits = 1, nsmall = 1)} with s.e. ",
    "{format(mad(v$y_pred), digits = 1, nsmall = 1)}\n",
    "Pr(Clinton win) = {format(mean(v$y_pred > 50), digits = 2, nsmall = 2)}"
  )
)
```

