---
title: "Regression and Other Stories: Coverage"
author: "Andrew Gelman, Jennifer Hill, Aki Vehtari"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---
Tidyverse version by Bill Behrman.

Coverage - Illustration of coverage of intervals. See Chapter 4 in
Regression and Other Stories.

-------------

```{r, message=FALSE}
# Packages
library(tidyverse)

# Parameters
  # Common code
file_common <- here::here("_common.R")

#===============================================================================

# Run common code
source(file_common)
```

## Data

```{r}
sim_mean <- 6
sim_sd <- 4
sim_n <- 100

sim <- 
  tibble(
    id = seq_len(sim_n),
    x = rnorm(sim_n, mean = sim_mean, sd = sim_sd),
    lower_95 = x + qnorm(0.025, mean = 0, sd = sim_sd),
    lower_50 = x + qnorm(0.25, mean = 0, sd = sim_sd),
    upper_50 = x + qnorm(0.75, mean = 0, sd = sim_sd),
    upper_95 = x + qnorm(0.975, mean = 0, sd = sim_sd)
  )
```

Coverage

```{r}
v <- 
  sim %>% 
  summarize(
    coverage_50 = mean(sim_mean >= lower_50 & sim_mean <= upper_50),
    coverage_95 = mean(sim_mean >= lower_95 & sim_mean <= upper_95)
  )

v
```

The 50% uncertainty interval contained the simulation population mean `r format(100 * v$coverage_50, digits = 0)`% of the time, and the 95% uncertainty interval contained the simulation population mean `r format(100 * v$coverage_95, digits = 0)`% of the time.

## Plot

```{r}
sim %>% 
  ggplot(aes(id)) +
  geom_hline(yintercept = sim_mean, color = "red") +
  geom_linerange(aes(ymin = lower_95, ymax = upper_95)) +
  geom_linerange(aes(ymin = lower_50, ymax = upper_50), size = 1) +
  geom_point(aes(y = x)) +
  labs(
    x = "Simulation",
    y = "Estimate, 50% and 95% uncertainty intervals"
  )
```

